### 2018-01-30 Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
add Radiocarbon abstract

---
title: "Data acquisition and analysis at NOSAMS"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution 
    footnote: Corresponding Author
  - name: Alan R. Gagnon
    affiliation: Woods Hole Oceanographic Institution 
  - name: Kalina Godispodinova
    affiliation: Woods Hole Oceanographic Institution 
  - name: Joshua Hlavenka
    affiliation: Woods Hole Oceanographic Institution 
  - name: Ann P. McNichol
    affiliation: Woods Hole Oceanographic Institution 
  - name: Mark L. Roberts 
    affiliation: Woods Hole Oceanographic Institution 
address:
  - code: Woods Hole Oceanographic Institution 
    address: WHOI/NOSAMS, Woods Hole, MA, 02543
abstract: |

Here we present progress on development of a hybrid sputter gas ion source and associated gas inlets.

We are focussing on two inlet technologies (compare and contrast) modified
hardware from a vg stable isotope mass spectrometer and an open split.

progress

application

goals

keywords: ion source, gas ion source, inlet, accelerator mass
spectrometry, radiocarbon output: rticles::elsevier_article ---