### 2018-04-03 Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
Final draft

---
title: "Development of a Hybrid Gas Ion Source and Gas Inlets at NOSAMS"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution 
    footnote: Corresponding Author
  - name: Alan R. Gagnon
    affiliation: Woods Hole Oceanographic Institution 
  - name: Joshua R. Burton
    affiliation: Woods Hole Oceanographic Institution 
  - name: Kalina D. Godispodinova
    affiliation: Woods Hole Oceanographic Institution 
  - name: Joshua D. Hlavenka
    affiliation: Woods Hole Oceanographic Institution 
  - name: Ann P. McNichol
    affiliation: Woods Hole Oceanographic Institution 
  - name: Mark L. Roberts 
    affiliation: Woods Hole Oceanographic Institution 
address:
  - code: Woods Hole Oceanographic Institution 
    address: WHOI/NOSAMS, Woods Hole, MA, 02543
abstract: |

  Here we present progress on development of gas inlet systems for a hybrid sputter gas ion source (HGIS). Our HGIS is a lightly modified gas-capable 134-sample MCSNICS source from NEC. The ion optics and cesium delivery have been upgraded following Roberts et al. [-@Roberts2010]. Gas delivery to the source is via a short glass capillary which crosses the source HV potential.

  We are developing two inlet technologies, one delivering gas to the source at low pressure using modified gas handling hardware from a VG stable isotope mass spectrometer and the other sampling gas at atmospheric pressure using an open split. The gas handling system designed for use with SIRMS is nearly ideal for our purposes, providing low volume, low pressure delivery of gas, and control of mixing ratios and delivery pressure using variable volumes. We will use the open split with large samples, focussing first on carbonate samples, hydrolysed and delivered using the gasbench system developed for use with the CFAMS microwave plasma GIS [@Roberts2013].

keywords: ion source, gas ion source, inlet, accelerator mass spectrometry, radiocarbon 
bibliography: RC-18HGISAbst.bib
csl: nuclear-inst-and-methods-in-physics-research-b.csl
output: rticles::elsevier_article 
---