### 2020-12-21 hgis_summary.Rmd
Add summary document

---
title: "USAMS Hybrid Source Summary"
author: "Brett Longworth"
date: "12/20/2020"
output: html_document
---

This document summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

# Source parameters

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 10uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.  Precision on modern
targets has been as good as 1% (SD of multiple measurements of a target) with
directly introduced bottle gas. Typical precision in tests of inlet systems has
been around 3%. Measured value of tank modern gas agrees with the accepted
value within error when normalizing to solid graphite standards.  The system
seems to have a constant blank contribution from the target and source of about
100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does not seem to add
significantly to this blank. Since the blank current is constant, the blank
contribution is inversely related to the sample current, ie. if the sample
current is 5uA, 2% of that will be from the blank and so on.

# Open split inlet

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uLs-1 flow.
Currently using 3m of 50um capillary with a bulkhead connector at the source
cage.  Precision of measurements of modern CO2 has been 3% at best, 10% at
worst, with better precision most recently. This is much less precise than the
expected 1%.  Accuracy is reasonable within the large errors on gas
measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
depending on sample current.

## Questions

Q: Background from the source/target is relatively stable at roughly 100nA or
1% of maximum CO2 current. Is the relative blank contribution proportional to
CO2 current?

A: Based on limited data, it's looking like yes, this follows a constant
contamination model.

Q: If the relative blank contribution is proportional to sample current, can we
correct for a varying sample current over a run (due to variable dillution,
pressure, etc.)?

A: If the sample current varies widely, this will likely require taking data in
shorter segments and blank correcting each segment using the mean current for
that segment.

Q: Why is the precision not as expected? Can we replicate the
literature/initial test precision?

A: Unknown. We can rerun a test with the best performing conditions (bottle gas
through capillary to source) to eliminate as many variables as possible.

## Development

2020-12-18 - Continue testing lower open split flows with leak checker, run
dead and live gas at lower flows. Spoke with Josh B. about double needle
setups.

2020-12-11 - Open split flow reduction. Test flow controller for helium, adjust
needle valves to allow lower CO2 flows. Measure blanks of target, helium, CO2
and dilluted CO2.

2020-12-04 - Measure blank contributions of target/source, helium, CO2, and
dilluted CO2. Install flow controller for helium.

2020-12-03 - Prepare and load new HGIS test wheel.

2020-11-17 - Diagnose issue with low currents, measure diluted and pure live
and dead gas.

2020-11-03 - Troubleshoot low currents, measure dilluted and pure live and dead
gas.

2020-10-15 - Troubleshoot low currents. Lots of tune and target position
testing.

2020-10-13 - Prepare and load new HGIS test wheel.

2020-10-06 - Dillution tests of live and dead gas. Test stability of currents
over time.

2020-09-29 - Starting up HGIS. Load sample wheel. Reterminate capillaries at
cage wall and open split.