### 2021-01-24 hgis_summary.Rmd
Update summary.

---
title: "USAMS Hybrid Source Summary"
author: "Brett Longworth"
date: "12/20/2020"
output: html_document
---

This document summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

# Source parameters

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 5uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.

Precision taken as the standard error of the mean of repeated runs of a sample is around 1%. Precision is typically a little worse than predicted by counting statistics. More work is needed to look at whether this reflects variability between samples.

Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples. Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.

The system seems to have a constant blank contribution from the target and
source of about 100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does
not seem to add significantly to this blank. Since the blank current is
constant, the blank contribution is inversely related to the sample current, ie.
if the sample current is 5uA, 2% of that will be from the blank and so on.

The above measurements are at optimum sample current of 6-10uA 12C-. Precision
and measured ratio both decrease as sample current drops below the optimum
range. The ratio effects may be related to the "constant current" blank
discussed above. More work is needed to see whether this is just a blank effect or whether there are other factors at low current.

# Open split inlet

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uL/min flow. A new open split with lower volume should have improved this, but testing is needed.

Currently, 250cm of 50um capillary defines flow rate from the open split into
the source, with a bulkhead connector at the source cage.  Precision of
measurements of modern CO2 has been better than 1% (std error of mean of
measurements of a target) for high current samples, with precision dropping off
as current decreases below 6uA or so. Accuracy is reasonable within the large
errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
to 5% of modern depending on sample current.

## Helium displacement from vials

First tests of displacing pure CO2 from a vial into the open split using helium
are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
over 2 hours. Currents do decrease with time due to dilution, but by supplying
higher than optimal CO2 flow to the source initially, stable currents can be
maintained for some time. Targets have shown "pulses" of higher current. Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.

## Questions

Q: Background from the source/target is relatively stable at roughly 100nA or
1% of maximum CO2 current. Is the relative blank contribution proportional to
CO2 current?

A: Based on limited data, it's looking like yes, this follows a constant
contamination model.

Q: If the relative blank contribution is proportional to sample current, can we
correct for a varying sample current over a run (due to variable dilution,
pressure, etc.)?

A: If the sample current varies widely, this will likely require taking data in
shorter segments and blank correcting each segment using the mean current for
that segment. We may be able to mitigate by setting capillary length such that
CO2 flow starts above optimal and drops to optimal by the end of the
acquisition.

Q: Does standard error of the mean as a sample error agree with between target variability?

A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.

## Development

2021-01-22 - Ran pure CO2 at different pressures to look at effect of current on precision and ratio. Repeated test for live and dead tank CO2.

2021-01-15 - Pure CO2 and displacement tests using vial and double needle. Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split. Followed this by switching to helium flow to vial at 100kPa (200ul/min). Repeated test with a second vial. Built second dual needle setup for use in vial filling (can also use Gilson).

2021-01-08 - Precision tests with direct CO2-source connection. Tested source pressure and current vs CO2 regulator pressure. Tested flow vs pressure with leak checker. Several capillary swaps. Acquired data at several flow rates for precision testing.

2021-01-07 - Set up for precision test using 7m x 50um capillary from CO2/he via Valco valve to cage wall.

2020-12-31 - Add CO2/He valve to vial test setup. VICI 4-way valve connected
with 1/16" ss tubing from CO2 and He regulators allows switching between gasses
delivered to a glass capillary leading to outer needle of dual needle. Leak
checking and capillary adjustments for good flow range at reasonable pressures.

2020-12-24 - Build dual needle setup. Supply displacement helium through outer
1/16" tubing, sample mixture through inner needle. Use VICI Luer adapter on top
of needle as open split with inserted capillary to source.

2020-12-18 - Continue testing lower open split flows with leak checker, run
dead and live gas at lower flows. Spoke with Josh B. about double needle
setups.

2020-12-11 - Open split flow reduction. Test flow controller for helium, adjust
needle valves to allow lower CO2 flows. Measure blanks of target, helium, CO2
and dilluted CO2.

2020-12-04 - Measure blank contributions of target/source, helium, CO2, and
dilluted CO2. Install flow controller for helium.

2020-12-03 - Prepare and load new HGIS test wheel.

2020-11-17 - Diagnose issue with low currents, measure diluted and pure live
and dead gas.

2020-11-03 - Troubleshoot low currents, measure dilluted and pure live and dead
gas.

2020-10-15 - Troubleshoot low currents. Lots of tune and target position
testing.

2020-10-13 - Prepare and load new HGIS test wheel.

2020-10-06 - Dillution tests of live and dead gas. Test stability of currents
over time.

2020-09-29 - Starting up HGIS. Load sample wheel. Reterminate capillaries at
cage wall and open split.