### 2021-02-22 hgis_summary.Rmd
Use include to use summary in README and summary doc.

---
title: "USAMS Hybrid Source Summary"
author: "Brett Longworth"
date: "`r Sys.Date()`"
output: html_document
---

This document summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

```{r child='summary.Rmd'}
```