### 2021-04-11 doc/SNEAP2021/SNEAP2021_Abstract.Rmd
edit for clarity and length

---
title: "Progress with the NOSAMS Hybrid Gas Ion Source"
author:
- Brett Longworth
- Josh Burton
- Josh Hlavenka
- Mark Roberts
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

A hybrid gas-sputter ion source produces carbon ions from CO~2~ gas for measurement of ^14/12^C via accelerator mass spectrometry. CO~2~ is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating CO~2~ and ionizing some of the carbon for extraction as in a solid sputter source. We are working to optimize ionization and develop interfaces that produce CO~2~ from samples and introduce it to the source. Early tests with CO~2~ standard gas gave way to introduction of CO~2~ diluted with helium carrier in a closed system, and most recent work introduces a CO~2~/He mixture to the source via an open split. Current development focuses on producing CO~2~ from carbonates (CaCO~3~) which is delivered to the open split and ion source using a flow of helium carrier gas. We will discuss efficiency, background carbon current, precision, and accuracy of the method.