### 2021-04-28 doc/SNEAP2021/SNEAP2021_figures.Rmd
figures notebook

---
title: "SNEAP2021 Figures"
output:
  html_document:
    df_print: paged
---

Figures for SNEAP talk

# Setup

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)
```

# Current vs flow

Good figure in first tests. Combine with data from dual bellows?

```{r, message = FALSE}

# factors for flow calc
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2

data <- get_hgis_data(here("data/USAMS040317R.txt")) %>% 
  filter(Pos == 13)

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data)) {
  if (data$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data$Run <- Run

# Load pressure data
pres <- read.csv(here("data/USAMS040317Pressure.csv"))
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data <- inner_join(data, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d.s <- data %>% 
  group_by(Run, Bottle, Source, flow) %>% 
  summarize(le12C = mean(le12C),
            he12C = mean(he12C),
            counts = sum(CntTotGT),
            interr = 1/sqrt(counts),
            C1412he = mean(X14.12he),
            C1412he_sd = sd(X14.12he),
            ferr = interr * C1412he,
            cor1412he = mean(cor1412he),
            eff = mean(eff))

# plot
filter(d.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}),
       y = bquote('12C-  ' *mu~ 'A')) +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

# Efficiency vs flow

# Dual bellows

Breakseal results for OX-I, C-1, OX-II

# Open split

## Dillution curve

From helium_displacement.Rmd

Maybe also show long dillution runs plotted by CO2 flow?

Yes. Calculate flow at time using concCO2

## Reproducibility of standards

Data from 1-08, 1-22, 1-29

## Blanks - blank values

compare dead gas, vial dead gas, C1, and TIRI-F

## Blanks - estimation of constant current contaminant

combine data from 12-04, 12-11, 12-18,

## Results for carbonates

Combine all runs 2-05, 3-05, 4-09, 4-16

show as agreement with consensus vs consensus value with sample names

Need normalized values, BC values, Errors, and rec_num or consensus value

```{r}
df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv)
```

```{r}
ggplot(df, aes(fm_consensus, mean - fm_consensus)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(df, aes(fm_consensus, diff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(df, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```