### 2021-04-30 doc/SNEAP2021/SNEAP2021_figures.Rmd
Redo secondaries figure. convert plots to svg

---
title: "SNEAP2021 Figures"
output:
  html_document:
    df_print: paged
---

Figures for SNEAP talk

# Setup

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)
library(patchwork)
```

# Current vs flow

Good figure in first tests. Combine with data from dual bellows?

```{r, message = FALSE}

# factors for flow calc
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2

data <- get_hgis_data(here("data/USAMS040317R.txt")) %>% 
  filter(Pos == 13)

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data)) {
  if (data$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data$Run <- Run

# Load pressure data
pres <- read.csv(here("data/USAMS040317Pressure.csv"))
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data <- inner_join(data, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d.s <- data %>% 
  group_by(Run, Bottle, Source, flow) %>% 
  summarize(le12C = mean(le12C),
            he12C = mean(he12C),
            counts = sum(CntTotGT),
            interr = 1/sqrt(counts),
            C1412he = mean(X14.12he),
            C1412he_sd = sd(X14.12he),
            ferr = interr * C1412he,
            cor1412he = mean(cor1412he),
            eff = mean(eff))

# plot
filter(d.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}),
       y = bquote('12C-  ' *mu~ 'A')) +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

Data from dual bellows tests

```{r}
read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
  mutate(flow = flowcalc(supply/10, 2E-5, 1.49E-5, 1.58)) %>% 
  ggplot(aes(flow, current)) +
  geom_smooth() + 
  geom_point() +
  labs(x = "Flow (ul/min)", y = "Source current (uA)")
```

# Efficiency vs flow

# Dual bellows

Breakseal results for OX-I, C-1, OX-II

# Open split

## Dillution curve

From helium_displacement.Rmd

Maybe also show long dillution runs plotted by CO2 flow?

Yes. Calculate flow at time using concCO2

```{r}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30 #30ul/min
         )
```

Sample current vs time

```{r}
# flow vs time subplot
flow_time <- data.frame(x = 0:125) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2")) +
  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "Gas flows to source",
       subtitle = "7mL vial, 244uL/min helium, 30ul/min to source",
       y = "Gas flow (uL/min)") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

# current vs time subplot
cur_time <- ggplot(data, aes(time, he12C)) +
  geom_smooth(span = .4, se=FALSE, color = "green") +
  geom_point() +
  xlim(0, 125) +
  labs(title = "Ion current",
       x = "Time (min)",
       y = "12C current (uA)")

flow_time / cur_time
```

### Expected CO2 ratio/flow

Plot volume of CO2 in sample vial during displacement by helium. It looks like optimal currents are maintained until about 2000s, which corresponds to about 2mL of CO2. Stable ratios until about 1.25 mL CO2. Can we back calculate flow to source if current falls off below ~2ul/min CO2?

Add flow at time to data and plot current vs flow. 30 ul/m is estimate of capillary flow based on past results.

```{r}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE) +
  geom_point() +
  labs(title = "Vial dillution current",
       subtitle = "250uL/min displacement, 30ul/min delivery",
       x = "CO2 flow (ul/m)",
       y = "12C current (uA)")
```

## Reproducibility of standards

Data from 1-08, 1-22, 1-29

## Blanks - blank values

compare dead gas, vial dead gas, C1, and TIRI-F

## Blanks - estimation of constant current contaminant

combine data from 12-04, 12-11, 12-18,

## Results for carbonates

Combine all runs 2-05, 3-05, 4-09, 4-16

show as agreement with consensus vs consensus value with sample names

Need normalized values, BC values, Errors, and rec_num or consensus value

```{r, message = FALSE}
df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
  filter(!is.na(rec_num)) %>% 
  mutate(fm_diff = mean - fm_consensus,
         fm_normdiff = normFm(mean, fm_consensus),
         fm_sigma = sigma(mean, fm_consensus, merr),
         fmbc_diff = fm_lbc - fm_consensus,
         fmbc_normdiff = normFm(fm_lbc, fm_consensus),
         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
```

```{r}
df %>% 
  filter(fm_normdiff < 100) %>% 
ggplot(aes(fm_consensus, fm_normdiff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
  geom_point() +
  ggtitle("Difference in measured Fm vs. consensus")
```

```{r}
df %>% 
  filter(fm_normdiff < 1) %>% 
ggplot(aes(fm_consensus, fm_normdiff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
  geom_point() +
  ggtitle("Difference in measured Fm vs. consensus")
```

```{r}
ggplot(df, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```

### difference from consensus

```{r}
secs <- df %>% 
  filter(fm_diff < .1) %>% 
ggplot(aes(fm_consensus, fm_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr)) +
  geom_point() +
  lims(x = c(0, 1.1),
       y = c(-.15, .15)) +
  ggtitle("Difference from consensus",
          subtitle = "Uncorrected") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())
```

```{r}
secs_lbc <- df %>% 
  filter(fm_diff < .1) %>% 
  mutate(mean_diff = mean(fmbc_diff)) %>% 
ggplot(aes(fm_consensus, fmbc_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  #geom_hline(yintercept = mean_diff) +
  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr)) +
  geom_point() +
  lims(x = c(0, 1.1),
       y = c(-.15, .15)) +
  labs(subtitle = "LBC-corrected")
```

```{r}
secs / secs_lbc
```