### 2021-06-10 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Move Rmd's to inst/Rmd

---
title: "HGIS 2021-05-28 Carbonates"
author: "Brett Longworth"
date: "6/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(HybridGIS)

options(digits = 4)
options(scipen = 3)
```

Data reduction and analysis for carbonate samples run on May 28, 2021. Includes C-1, C-2, and test shell carbonates from Simon Pendleton.

```{r}

#Raw AMS data results file path
file <- here("data/USAMS052621R.txt")

# Date samples run (needed to separate other experiments with this wheel/targets)
date <- as.Date("2021-05-28")

# Positions of normalizing standards
standards <- c(5, 8, 13, 17)

# Positions of blanks for blank correction
blanks <- c(6, 9, 11)
```

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors.

## raw results

```{r}
df <- get_hgis_data(file, date, standards)
write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11))

write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
```

## Results

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, fm_corr, sig_fm_corr) %>% 
  arrange(Sample.Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
plot_hgis_summary()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates() %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(Sample.Name)

knitr::kable(cons)
```