### 2021-06-10 inst/rmd/HGIS_small_samples.Rmd

---
title: "HGIS small(er) samples"
author: "Brett Longworth"
date: "6/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# Load libraries and functions
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)
```

## 2021-06-04 Small(er) sample tests

Tested 5, 10, and 15 mg of C-2 carbonate in vials backfilled with Helium.

### Method

1. Weigh sample into vial
2. Evacuate and acidify as normal with gilson
3. Backfill to 1 ATM with He-filled syringe 2h after acidification.

Allowed vacuum in vial to pull in He from syringe, then added 1 additional mL He. Should calculate volume CO2 in vial and measure He added to check that things make sense.

Used same capillaries as usual, but decreased He displacement flow to 140 ul/min. Source capillary flow assumed to be around 30ul/min.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-2_1", 1082, 6.03,
                     2, "C-2_2", 1082, 11.10,
                     3, "C-2_3", 1082, 15.52
                     )

write_csv(carb_data, here("data/carb_data_2021-06-04.csv"))
carb_data
```

### Data

```{r}
df <- get_hgis_data(here("data/USAMS052621R.txt"), as.Date("2021-06-04")) #, standards)
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

### Results

Note: no blanks run, so need to use average blank if LBC needed.

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
  arrange(Sample.Name)
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(Sample.Name, norm_ratio)) +
    geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
                    size = .5) + 
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
      labs(x = NULL,
           y = "Fraction Modern")
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates()
```

## Agreement with consensus

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
  mutate(Fm_diff = norm_ratio - fm_consensus,
         sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
  arrange(Sample.Name)
```