### 2021-06-10 inst/rmd/RC2018Figures.Rmd

---
title: "Figures for Radiocarbon 2018 Poster"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r libraries, warning=FALSE, message=FALSE}
library(amstools)
library(tidyverse)
library(ggplot2)
```

## Calculated flow

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

The following functions calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{m}$ and pressure in kPa.

```{r}
# Poiseuille's; convert delta pressure in Pa to flow in m^3/s
prestoflow <- function(dp, r, u, l) {
  # r = radius of capillary
  # l = length of capillary
  # u = viscosity kg/m
  flow <- dp * pi * r^4 / 8 / u / l
  flow
}

# convert gauge kPa to uL/min
flowcalc <- function(pres, ...) {
  dppa <- pres * 1E3 #convert kPa to Pa
  flowcms <- prestoflow(dppa, ...)
  flowuls <- flowcms * 1E6 * 1E3
  flowulm <- flowuls * 60
  flowulm
}
```

```{r}
# Efficiency functions
# convert L CO2 to atoms C
lCO2toatC <- function (vol) {
  molC <- vol * 1/22.4 # 22.4 L of gas per mole at stp
  atC <- molC * 6.022E23
  atC
}

# convert current in A to atoms C/sec
CcurtoatC <- function(cur, cs){
  at12C <- cur * 6.25E18 # electrons per coulomb
  at12C <- at12C / cs # divide by charge state
  at12C
}

# calculate efficiency from ml/m CO2 and uA C12
hgis_eff <- function(ulm, uA, cs = 3) {

  # if flow is zero or less, efficiency doesn't make sense
  ulm <- ifelse(ulm <= 0, NA, ulm)

  vol <- ulm * 1E-6 # conv to L
  cur <- uA * 1E-6 # conv to A
  atCin <- lCO2toatC(vol) / 60 # gives atoms/s
  atC12out <- CcurtoatC(cur, cs)
  atCout <- atC12out / .99 # 99% of C is 12C, we need total C
  atC12out / atCin
}
```

## USAMS032817

```{r}
data <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS032817R.txt") %>%
  mungeResfile() 
```

### Currents

For the first target, we turned up the gas supply pressure while measuring
source current at the high energy FC. These currents are roughly equivalent to
le12C.

* TODO: Blank current subtraction

Load and massage the data...

```{r}
cur <- read.delim("../Data/1sttestcurrents.csv") %>%
  mutate(flow = flowcalc(regp, r, u, x),
         eff = hgis_eff(flow, he12c, cs = 1)) %>% # cs 1 to estimate LE current
  filter(t > 5) 
```

```{r}
ggplot(cur, aes(regp, he12c)) + geom_line() + geom_point()
ggplot(cur, aes(flow, he12c)) + geom_line() + geom_point()
```

Currents were very stable after an initial burn in of ~10m. Initial current
started at ~12uA without gas on target and decayed (roughly) exponentially to a
minimum around 1.2uA.

```{r}
data %>% filter(Pos %in% c(11,12)) %>%
       	ggplot(aes(ts, he12C, color = Pos)) +
       	geom_line() + geom_point() + ylim(c(0,8))
```

### Efficiency

Using efficiency functions defined above. Efficiency drops with increasing flow
as currents go up.

```{r, warning = FALSE}
ggplot(cur, aes(flow, eff)) + geom_line() + geom_point()
```

### Ratios

Ratios for live and dead CO2 were recorded in 180 blocks of 10s each. Solid AA
and OX-I were run with 10 blocks of 180s.

#### Summary table

```{r}
data.s <- data %>% group_by(Pos) %>% summarize(c1412.m = mean(X14.12he),
                                         c1412.s = sd(X14.12he),
                                         c1412.sdr = c1412.s/c1412.m,
                                         c1412.se = se(X14.12he),
                                         c1412.ser = c1412.se/c1412.m,
                                         counts = sum(CntTotGT),
                                         interr = 1/sqrt(counts))
knitr::kable(data.s, digits = 4)
```

#### Plots

Boxplots of modern and dead samples. Ratios were essentially stable over time.
Positions < 10 are solid samples run for comparison.

Modern samples.

```{r}
data %>% filter(X14.12he > .4) %>%
       	ggplot(aes(Pos, X14.12he, color = Num)) + 
        geom_boxplot()
```

Dead gas and alpha aesar.

```{r}
data %>% filter(X14.12he < .4) %>% ggplot(aes(Pos, X14.12he, color = Num)) + 
        geom_boxplot()
```

```{r, message = FALSE}
data2 <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS040317R.txt") %>%
  filter(Pos == 13) %>%
  mungeResfile()

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data2)) {
  if (data2$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data2$Run <- Run

# Load pressure data
pres <- read.csv("../Data/USAMS040317Pressure.csv")
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data2 <- inner_join(data2, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d2.s <- data2 %>% group_by(Run, Bottle, Source, flow) %>% summarize(le12C = mean(le12C),
                                              he12C = mean(he12C),
                                              counts = sum(CntTotGT),
                                              interr = 1/sqrt(counts),
                                              C1412he = mean(X14.12he),
                                              ferr = interr * C1412he,
                                              cor1412he = mean(cor1412he),
                                              eff = mean(eff))
```

# Current vs Flow

```{r, message = FALSE, warning = FALSE}
ggplot(d2.s, aes(Bottle, he12C)) +
  geom_smooth() +
  geom_point() 
filter(d2.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = "CO2 flow rate (uL/min)", y = "12C- current (uA)") +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

# Efficiency vs Flow

```{r, message = FALSE, warning = FALSE}
ggplot(d2.s, aes(flow, eff*100)) + 
  geom_smooth() +
  geom_point() +
  labs(x = "CO2 flow rate (uL/min)", y = "Measurement efficiency (%)") +
  xlim(0,15.5)
```

# Modern Performance

# Blank