### 2021-06-16 inst/rmd/carbonate_reanalysis.Rmd
fix stds, don't remove outliers, better summaries

---
title: "Carbonate reanalysis"
author: "Brett Longworth"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reanalysis of carbonate data with most current data reduction method.

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)
```

```{r}
wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121")
files <- here("data", paste0("USAMS", wheels, "R.txt"))
dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11"))
stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL)
blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL)

results <- pmap(list(files, dates, stds, blanks), reduce_hgis, remove_outliers = FALSE)
raw_df <- map_dfr(results, 1)
results_df <- map_dfr(results, 2)
```

```{r}
results_df %>% 
  filter(!is.na(rec_num),
         sample_type == "B") %>% 
  select(sample_name, rec_num, wheel, sample_type, fm_consensus, fm_corr, sig_fm_corr) %>% 
  arrange(pos)
```

```{r}
results_df %>% 
  filter(!is.na(rec_num),
         !is.na(sig_fm_corr)) %>% 
  compare_replicates() %>% 
  arrange(rec_num)
```