### 2021-06-16 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Add comparison to graphite

---
title: "HGIS 2021-05-28 Carbonates"
author: "Brett Longworth"
date: "6/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(HybridGIS)

options(digits = 4)
options(scipen = 3)
```

Data reduction and analysis for carbonate samples run on May 28, 2021. Includes C-1, C-2, and test shell carbonates from Simon Pendleton.

```{r}

#Raw AMS data results file path
file <- here("data/USAMS052621R.txt")

# Date samples run (needed to separate other experiments with this wheel/targets)
date <- as.Date("2021-05-28")

# Positions of normalizing standards
standards <- c(5, 8, 13, 17)

# Positions of blanks for blank correction
blanks <- c(6, 9, 11)
```

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
- Outliers are NOT dropped from the analysis.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors

## raw results

```{r}
df <- process_hgis_results(file, date, standards)
#write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
#write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(pos %in% 1:4)) %>% 
  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
  labs(title = "Sample ratio vs. time",
       y = "Fraction modern")
```

```{r}
df %>% 
  filter(!(pos %in% 1:4)) %>% 
  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
  labs(title = "Sample current vs time",
       x = "Time (s)",
       y = "Current (uA)")
```

```{r}
df_sum <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11))

# write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
# write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
```

## Results

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, 
         `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
  arrange(Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
plot_hgis_summary()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
  compare_replicates() %>% 
  select(Name, `Mean current (uA)` = he12C_mean, 
         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- df_sum %>% 
  filter(as.numeric(pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(sample_name) %>% 
  select(Name = sample_name, 
         #`Mean current (uA)` = he12C_mean, 
         #`SD of current` = he12C_sd, 
         `Consensus Fm` = fm_consensus,
         Fm = fm_corr, 
         `Fm error` = sig_fm_corr,
         `Difference` = Fm_diff,
         Sigma = sigma)

knitr::kable(cons)
```

## Agreement with graphite

Agreement of Haiti core samples. Graphite samples were pretreated with HCl to leach 30% of the sample. HGIS samples not pretreated.

Load Simon's data and compare

```{r}
haiti <- filter(df_sum, str_detect(sample_name, "HATC1")) %>% 
  mutate(rec_num = case_when(pos == 14 ~ 171989,
                             pos == 15 ~ 171995,
                             pos == 16 ~ 171999))

graphite_results <- getRecOS(c(171999,171995,171989)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)

combined_results <- haiti %>% 
  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
  left_join(graphite_results,
            by = "rec_num", 
            suffix = c("_hgis", "_graphite")) %>% 
  mutate(fm_diff = fm_corr_graphite - fm_corr_hgis,
         fm_mean = mean(c(fm_corr_hgis, fm_corr_graphite))) %>% 
  pivot_longer(c(fm_corr_graphite, sig_fm_corr_graphite, 
                 fm_corr_hgis, sig_fm_corr_hgis), 
               names_to = c(".value", "method"),
               names_pattern = "(.*)_(.*)") %>% 
  mutate(fm_diff = fm_corr - fm_mean)

combined_results %>% 
  select(sample_name, rec_num, method, fm_corr, sig_fm_corr) %>% 
knitr::kable()
```

Difference plot

```{r}
combined_results %>% 
  ggplot(aes(fm_corr, fm_diff, color = method, shape = method)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr)) +
                  #position = position_dodge2(width = 0.05, padding = 0.1)) +
  ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm") + theme_classic()
```