### 2021-06-20 doc/data_meeting_2021-06-22/data_meeting_2021-06-22.Rmd
Add data meeting talk

---
title: "HGIS Carbonate Results"
subtitle: "NOSAMS Data Meeting"
author: 
  - "Brett Longworth"
date: '2021-06-22'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

```{r}
library(tidyverse)
library(HybridGIS)
library(here)
library(gt)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

## Method

.pull-left[
1. Acidify
  - ~30mg Carbonate produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials
]

.pull-right[![HGIS Figure](images/HGIS_open_split_drawing.svg)]
---

## Data reduction

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

---

## Results: Agreement with Consensus

```{r}
# Read data. Don't use first carbonate test or smaller samples.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
cons_df <- results_df %>% 
  filter(!is.na(sig_fm_corr),
         wheel != "USAMS061121") %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus()
```

```{r}
plot_hgis_consensus(cons_df) +
  scale_color_continuous(name = "12C- (μA)")
```
---

.left-column[
## Results: Method precision and accuracy
]

.right-column[
```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, Fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```
<br>
<br>

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, Fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = Fm_diff_mean,
         `SD of mean difference` = Fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  #pivot_longer(everything()) %>% 
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```
]

---

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = Fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  #geom_boxplot(outlier.shape = NA) +
  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```