### 2021-06-24 R/HGISFirstTest.Rmd
reorganize

---
title: "First Hybrid GIS Tests"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
date: "03-28-2017"
---

First tests of the USAMS SNICS-134 Hybrid gas ion source. Where not specified, flows are given in $\frac{\mu L}{m}$ and pressure in kPa, and ion currents are in uA of 12C.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)

theme_set(theme_bw())
options(digits = 3)
```

# Flow rates

## Calculated flow

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

The following functions calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{m}$ and pressure in kPa.

```{r}
# Poiseuille's; convert delta pressure in Pa to flow in m^3/s
prestoflow <- function(dp, r, u, l) {
  # r = radius of capillary
  # l = length of capillary
  # u = viscosity kg/m
  flow <- dp * pi * r^4 / 8 / u / l
  flow
}

# convert gauge kPa to uL/min
flowcalc <- function(pres, ...) {
  dppa <- pres * 1E3 #convert kPa to Pa
  flowcms <- prestoflow(dppa, ...)
  flowuls <- flowcms * 1E6 * 1E3
  flowulm <- flowuls * 60
  flowulm
}
```

Initial tests used a 0.050mm x 8.2m deactivated silica capillary. For this
capillary, parameters for Poiseuille's equation are as follows:

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2
factor <- flowcalc(1, r, u, x)
```

$\Delta P$ is regulator pressure

R is `r r*1E6` $\mu m$

$\mu$ is `r u*1E6` $\mu Pa\cdot s$

x is `r x`m

Given these,

Q($\frac{\mu L}{m}$) = `r factor` x P (kPa)

```{r}
flow <- function(f) {
  flowcalc(f, r, u, x)
}
ggplot(data = data.frame(pressure = 0, flow = 0),
       mapping = aes(x = pressure, y = flow)) +
  stat_function(fun = flow) +
  xlim(0,400) +
  ggtitle("Theoretical flow for given pressure",
          subtitle = "8.2m x 50um capillary")
```

## Measured flow

Kalina and I measured flow at the end of the capillary using two methods;
direct measurement with a bubble flow meter and as a helium leak rate with the
leak detector.

### Bubble meter

Kalina made measurements with both helium and CO2. The measurements below are
with helium and converted to CO2 by the ratio of their viscosities.

```{r}
bflow <- read.csv(here("data/bubbleflow.csv")) %>%
  mutate(heflow = volume/time,
         flow = heflow * 1.49E-5/1.90E-5, # co2/He viscosity
         pressure = pressure - 14.7, # subract 1 Atm for delta P
         pressure = pressure * 6.8948, # convert psi to kPa
         calcflow = flowcalc(pressure, r, u, x)) # calculate flow with Poiseuille's

ggplot(bflow, aes(pressure, flow)) + 
  stat_function(fun = flow, color = "blue") +
  geom_point() +
  geom_line() +
  xlim(0,400) + ylim(0,80) +
  ggtitle("Flow vs pressure",
          subtitle = "measured by gauge and bubble flow meter")

ggplot(bflow, aes(calcflow, flow)) + 
  geom_point() +
  geom_line() +
  ggtitle("Measured (bubble meter) vs predicted flow")
```

### Leak detector

#### 2017-10-17?

Measured leak rate using Helium leak detector. Ran capillary to KF adapter
flange. Pressures in kPa, leak rates in $\frac{Pa\cdot m^3}{s}$. Leak detector
switched to gross leak mode between 150 kPa and 200 kPa.

```{r}
ldflow <- read.csv(here("data/HeLeakCapillaryFlow.csv"), skip = 3)

ldflow <- ldflow %>% 
  mutate(sccm = Leakrate / 0.0016, # convert to sccm
         co2sccm = sccm * 1.49E-5/1.90E-5, # co2/He viscosity
         flow = co2sccm * 1E3, # convert to ul/m
         calcflow = flowcalc(Regulator, r, u, x)) # calculate flow with poiseulle's

ggplot(ldflow, aes(Regulator, flow)) + 
  stat_function(fun = flow, color = "blue") +
  geom_point() +
  geom_line() +
  ggtitle("Flow (measured by leak detector) vs pressure")

ggplot(ldflow, aes(calcflow, flow)) + 
  geom_point() +
  geom_line() +
  ggtitle("Measured (by leak detector) vs predicted flow")

```

## Efficiency functions

These functions convert volume of CO2 and coulombs of charge to atoms of C. The
ratio of these is the measurement efficiency of a gas ion source system. These are used for later calculations of efficiency given flow rate and ion current.

```{r}
# convert L CO2 to atoms C
lCO2toatC <- function (vol) {
  molC <- vol * 1/22.4 # 22.4 L of gas per mole at stp
  atC <- molC * 6.022E23
  atC
}

# convert current in A to atoms C/sec
CcurtoatC <- function(cur, cs) {
  at12C <- cur * 6.25E18 # electrons per coulomb
  at12C <- at12C / cs # divide by charge state
  at12C
}

# calculate efficiency from ml/m CO2 and uA C12
hgis_eff <- function(ulm, uA, cs = 3) {

  # if flow is zero or less, efficiency doesn't make sense
  ulm <- ifelse(ulm <= 0, NA, ulm)

  vol <- ulm * 1E-6 # conv to L
  cur <- uA * 1E-6 # conv to A
  atCin <- lCO2toatC(vol) / 60 # gives atoms/s
  atC12out <- CcurtoatC(cur, cs)
  atCout <- atC12out / .99 # 99% of C is 12C, we need total C
  atC12out / atCin
}
```

# Test Wheels

Two days of testing to date. Same wheel both days. Wheel had 5 AA, 4 OX-I, and
10 gas targets.

## USAMS032817

### Currents

For the first target, we turned up the gas supply pressure while measuring
source current at the high energy FC. These currents are roughly equivalent to
le12C (~30% transmission @ 3+ charge state).

Load and massage the data...

```{r}
cur <- read_tsv(here("data/2017-03-28Currents.tsv")) %>%
  mutate(flow = flowcalc(regp, r, u, x),
         eff = hgis_eff(flow, he12c, cs = 1)) %>% # cs 1 to estimate LE current
  filter(t > 5) 
```

```{r}
ggplot(cur, aes(regp, he12c)) +
  geom_line() +
  geom_point() +
  ggtitle("Current vs Pressure",
          subtitle = "2017-03-28")

ggplot(cur, aes(flow, he12c)) +
  geom_line() +
  geom_point() +
  ggtitle("Current vs Flow",
          subtitle = "2017-03-28")
```

```{r}
data <- get_hgis_data(here("data/USAMS032817R.txt")) %>%
           mutate(Num = ifelse(Pos == 11, "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S"])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
# data <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS032817R.txt") %>%
#   mungeResfile() 
```

Currents were very stable after an initial burn in of ~10m. Initial current
started at ~12uA without gas on target and decayed (roughly) exponentially to a
minimum around 1.2uA.

```{r}
data %>% 
  filter(Pos %in% c(11,12)) %>%
  ggplot(aes(ts, he12C, color = Pos)) +
   	geom_line() + geom_point() + ylim(c(0,8))
```

### Efficiency

Using efficiency functions defined above. Efficiency drops with increasing flow
as currents go up.

```{r, warning = FALSE}
ggplot(cur, aes(flow, eff)) + geom_line() + geom_point()
```

### Ratios

Ratios for live and dead CO2 were recorded in 180 blocks of 10s each. Solid AA
and OX-I were run with 10 blocks of 180s. These results are summarized and roughly normalized to solid OX-I below.

#### Summary table

```{r}
sum_hgis(data)
```

#### Plots

Boxplots of modern and dead samples. Ratios were essentially stable over time.
Positions < 10 are solid samples run for comparison.

Modern samples.

```{r}

data %>% 
  filter(X14.12he > .4,
         Pos %in% c(7, 11)) %>%
  mutate(type = ifelse(Pos == 7, "solid", "gas")) %>%
       	ggplot(aes(type, X14.12he)) + 
        geom_boxplot() +
        labs(x = "Sample type", y = "Fraction modern")
```

Dead gas and alpha aesar.

```{r}
data %>% filter(X14.12he < .02,
                X14.12he > 0.0005,
                Pos %in% c(3, 12)) %>%
  mutate(type = ifelse(Pos == 3, "solid", "gas")) %>%
           ggplot(aes(type, X14.12he)) + 
        geom_boxplot() +
        labs(x = "Sample type", y = "Fraction modern")

```

## USAMS040317

The same target wheel was loaded again. Karl and I experimented with the source
tune and target position to try and get more current. We were able to increase
the currents by about 1/3 by moving the target closer to the immersion lens.
Flow rates were lower for a given current. After tuning, I looked at flow vs.
ratio and current. The gas used was the modern standard, but the flask may not
have been fully flushed.

```{r, message = FALSE}
data <- get_hgis_data(here("data/USAMS040317R.txt")) %>% 
  filter(Pos == 13)

sum_hgis(data)

# data <- readResfile(here("data/USAMS040317R.txt")) %>%
#   filter(Pos == 13) %>%
#   mungeResfile()

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data)) {
  if (data$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data$Run <- Run

# Load pressure data
pres <- read.csv(here("data/USAMS040317Pressure.csv"))
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data <- inner_join(data, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d.s <- data %>% group_by(Run, Bottle, Source, flow) %>% summarize(le12C = mean(le12C),
                                              he12C = mean(he12C),
                                              counts = sum(CntTotGT),
                                              interr = 1/sqrt(counts),
                                              C1412he = mean(X14.12he),
                                              C1412he_sd = sd(X14.12he),
                                              ferr = interr * C1412he,
                                              cor1412he = mean(cor1412he),
                                              eff = mean(eff))
d.s
```

#### Gas flow vs source pressure

```{r}
ggplot(d.s, aes(Bottle, Source)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(flow, Source)) + 
  geom_smooth() +
  geom_point() 
```

#### Currents

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(Bottle, he12C)) +
  geom_smooth() +
  geom_point() 
filter(d.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}),
       y = bquote('12C-  ' *mu~ 'A')) +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

#### Ratio vs current

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(he12C, C1412he)) + 
  geom_linerange(aes(ymin = C1412he - ferr, ymax = C1412he + ferr)) +
  geom_smooth() +
  geom_point() 
```

#### Efficiency

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(Bottle, eff)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(Source, eff)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(flow, eff*100)) + 
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}), y = "Measurement efficiency  %") +
  xlim(0,15.5)
```