### 2021-06-24 R/HGISFlowCalcs.Rmd

---
title: "HGIS flow calcs"
output:
  html_document: default
  html_notebook: default
---

```{r}
library(here)
library(HybridGIS)
```

# Calculated Flow rates

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

  The functions loaded above calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{min}$ and pressure in kPa.

## Initial NOSAMS parameters

Initial tests used a 0.050mm x 8.2m deactivated silica caplillary. For this
capillary, parameters for Poiseuille's equation are as follows:

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2
flowcalc(100, r, u, x)
```

## Parameters for published hybrid GIS systems

Calculate flow for Xu et al 2006...

```{r}
flowcalc(1.500, 0.000125, u, 1.25)
```
Calculate flow for ETH...

```{r}
flowcalc(15.6, 3E-5, u, 2)
flowcalc(31.3, 3E-5, u, 2)

```
Calculate flow for CAMS...

```{r}
flowcalc(42, 7.5E-5, u, 6)

```
Calculate flow for Oxford...

```{r}
flowcalc(200, 3E-5, u, 1.25)

```

## Parameters for next USAMS tests

### Short capillary at 1 atm

40um capillary

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 3.5
flowcalc(100, r, u, x)
```
```{r}
r <- 2E-5
u <- 1.49E-5
x <- 2
flowcalc(100, r, u, x)
```
50um capillary
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 5
flowcalc(101, r, u, x)
```
75um capillary
```{r}
r <- 3.75E-5
u <- 1.49E-5
x <- 2.8
flowcalc(101, r, u, x)
```

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 7
flowcalc(100, r, u, x)
```
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 3
flowcalc(20, r, u, x)
```
### Low pressure 15 mbar

100um capillary

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(1.5, r, u, x)
```

### Medium pressure 150mbar

75um capillary

```{r}
r <- 3.75E-5
u <- 1.49E-5
x <- 4
flowcalc(15, r, u, x)
```

100um capillary

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 2.0
flowcalc(5, r, u, x)
```
Medium pressure 150mbar 100um x 1.58m PEEK capillary
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(20, r, u, x)
```

10x dillution
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(15, r, u, x)
```

Theoretical flows with pressure adjusted for best current.

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(60, r, u, x)
```
Actual performance of "100um"" x 1.58m PEEK capillary *IF* flow is ~10uL/min. Looks like
true diameter after crimping is around 40um.

```{r}
r <- 2.0E-5
u <- 1.49E-5
x <- 1.58
flowcalc(60, r, u, x)
```

# CO2 quantities

Calculate CO2 needed for an analysis at given supply pressure.