### 2021-06-24 README.Rmd
Update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HybridGIS

<!-- badges: start -->
<!-- badges: end -->

The HybridGIS project contains analysis scripts and data for the NOSAMS hybrid gas ion source.

## Setting up

This project uses the amstools and hgis packages, both available from github. They may be installed as follows:

```{r, message=FALSE}
remotes::install_github(c("blongworth/amstools", "blongworth/hgis"))
```

Running `make` will download the NOSAMS results files if a connection to the NOSAMS file server is available. This currently uses an absolute path to the server of `/mnt/shared/USAMS/Results`. This should be converted to a URL, but until then, modify `RES_DIR` in `Makefile` to match it's location on your system. `make` will eventually run the analysis scripts as well.

## Functions

The `amstools` package provides general functions for ams data retrieval and reduction, and the `hgis` package has functions for data reduction, flow calculation, efficiency, data import and plotting of HGIS data.

```{r example, message=FALSE}
library(amstools)
library(hgis)
```

# Project Summary

The information below summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

```{r child='summary.Rmd'}
```

# Presentations

Slides for a presentation given at SNEAP 2021:

### [SNEAP 2021 Presentation](doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.html)