### 2021-06-27 R/smaller_samples.Rmd
Add data and analysis for samples run 2021-06-25.

---
title: "Small(er) Samples"
author: "Brett Longworth"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(amstools)
library(hgis)

options(digits = 3)
options(scipen = 4)

df <- read_csv(here("data_analysed/carb_all_results.csv")) %>% 
  compare_consensus() %>% 
  mutate(c12ua = he12C * 1E6)
samples <- read_csv(here("data/carb_data.csv")) %>% 
  select(wheel, pos, mass, he_added)
df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
  filter(wheel != "USAMS020521")
```

```{r}
df %>% 
ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +
  geom_point()
```

```{r}
df %>% 
  filter(wheel == "USAMS062521") %>% 
ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(mass, fm_diff, color = wheel)) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(mass, fm_diff, color = as.factor(rec_num))) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(c12ua, abs(fm_diff))) +
  geom_smooth() +
  geom_point(aes(color = wheel))
```

```{r}
df %>% 
  filter(sample_type == "S") %>% 
ggplot(aes(c12ua, abs(fm_diff))) +
  geom_point(aes(color = wheel))
```