### 2021-07-05 R/HGIS_pendleton_all.Rmd
Add new Haiti data. Comparison looks great!

---
title: "HGIS Pendleton Carbonates"
author: "Brett Longworth"
date: "2021-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(hgis)

options(digits = 4)
options(scipen = 3)
```

Data analysis for Haiti core carbonates from Simon Pendleton. Data reduction is handled by `carbonate_samples.R` using the method and functions from the `hgis` package.

- Using standards in positions XXX for normalization.
- Using blanks in positions XXX for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
- Outliers are NOT dropped from the analysis.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors

## Load results

```{r}
data <- readRDS(here("data_analysed/carb_all.Rdata"))
sample_data <- read_csv(here("data/carb_data.csv")) #%>% 
  #select(wheel, pos, mass, he_added)

raw <- map_dfr(data, 1) %>%  
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4)

results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)
```

#### Currents and ratios vs time.

```{r}
raw %>% 
  plot_hgis_time(norm_ratio, outlier = ok_calc) +
  labs(title = "Sample ratio vs. time",
       y = "Fraction modern")
```

```{r}
raw %>% 
  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
  labs(title = "Sample current vs time",
       x = "Time (s)",
       y = "Current (uA)")
```

## Results

```{r}
results %>% 
  mutate(he12C = he12C * 1E6) %>% 
  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, 
         `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
  arrange(Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
plot_hgis_summary(results)
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
results %>% 
  compare_replicates() %>% 
  select(Name, `Mean current (uA)` = he12C_mean, 
         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- results %>% 
  filter(!is.na(fm_consensus)) %>% 
  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(sample_name) %>% 
  select(Name = sample_name, 
         #`Mean current (uA)` = he12C_mean, 
         #`SD of current` = he12C_sd, 
         `Consensus Fm` = fm_consensus,
         Fm = fm_corr, 
         `Fm error` = sig_fm_corr,
         `Difference` = Fm_diff,
         Sigma = sigma)

knitr::kable(cons)
```

## Agreement with graphite

Agreement of Haiti core samples. Graphite samples were pretreated with HCl to leach 30% of the sample. HGIS samples not pretreated.

Load Simon's data and compare

```{r}
graphite_results <- getRecOS(results$rec_num) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)

combined_results <- results %>% 
  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
  inner_join(graphite_results,
            by = "rec_num", 
            suffix = c("_hgis", "_graphite")) %>% 
  mutate(fm_diff = fm_corr_graphite - fm_corr_hgis,
         fm_mean = mean(c(fm_corr_hgis, fm_corr_graphite))) %>% 
  pivot_longer(c(fm_corr_graphite, sig_fm_corr_graphite, 
                 fm_corr_hgis, sig_fm_corr_hgis), 
               names_to = c(".value", "method"),
               names_pattern = "(.*)_(.*)") %>% 
  mutate(fm_diff = fm_corr - fm_mean)

combined_results %>% 
  select(sample_name, rec_num, method, fm_corr, sig_fm_corr) %>% 
  arrange(rec_num) %>% 
knitr::kable()
```

Difference plot

```{r}
combined_results %>% 
  ggplot(aes(fm_corr, fm_diff, color = method, shape = method)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1,
                  position = position_dodge2(width = 0.02)) +
  ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm") + theme_classic()
```