### 2021-07-26 doc/method_validation/method_validation.Rmd
Add work on method validation and error propagation.

---
title: "Error propagation error"
output: html_notebook
---

Trying to find the source of differences between MR's excel error prop and mine in R.

```{r}
library(tidyverse)
library(hgis)
library(readxl)
library(here)
library(errors)
```

Load results from R data reduction
```{r}
results <- readRDS(here("data_analysed/carb_all.Rdata"))
df <- map_dfr(results[10], 2)
View(df)
```

Load MR's results. I think we'd have to grab individual cells to do this.

```{r}
dfm <- read_excel(here("doc/method_validation/HGIS\ Results.xlsx"))
```

Compare raw ratios and errors

Looking at the data, the difference is in the 2 methods for calculating the error in the Fm of normalizing standards.

MR's method:

```{r}
stds <- df %>% 
  ungroup() %>% 
  filter(sample_type == "S") %>% 
  select(pos, corr_14_12, int_err, ext_err, max_err)

stds
```

```{r}
summarize(stds, ratio_std = mean(corr_14_12),
            sig_ratio_std = sqrt(sum(max_err^2))/n())
```

BL's method

```{r}
summarize(stds, ratio_std = mean(corr_14_12),
            sig_ratio_std = amstools::se(corr_14_12))
```

Propagated error of averaging using the `errors` package.

```{r}
stds <- stds %>% 
  mutate(corr_14_12_err = set_errors(corr_14_12, max_err))

ratio_std <- stds%>% 
  summarize(ratio_std = mean(corr_14_12_err)) %>% 
  pull(ratio_std)
```

Trying to get Fm using errors package propagation

```{r}
fm_std <- set_errors(1.03098, 0.0006)

df <- df %>% 
  mutate(corr_14_12_err = set_errors(corr_14_12, max_err),
         norm_ratio_err = norm_gas(corr_14_12_err, ratio_std, fm_std))

df %>% select(pos, sample_name, norm_ratio, sig_norm_ratio, norm_ratio_err)

```