### 2021-07-26 doc/method_validation/normalization error.Rmd

---
title: "Normalization Error"
author: "Brett Longworth"
date: "7/14/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mark's method

Unknowns are normalized to the mean ratio of standards and then converted to Fm using the Fm of the normalizing standard:

$$ R_{norm} = \frac{R_{sample}}{R_{std}}$$

$$Fm_{sample} = R_{sample}\cdot Fm_{std}$$

With error propagation in 2 steps:

$$ \sigma_{R_{norm}} = R_{norm} \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2}}$$

$$ \sigma_{Fm_{sample}} = Fm_{sample} \sqrt{\frac{\sigma_{R_{norm}}^2}{R_{norm}^2} + \frac{\sigma_{Fm_{std}}^2}{Fm_{std}^2}} $$

## Normalize targets using mean of standards and propagate errors

Unknowns are normalized using the mean measured value of gas standards using this equation:

$$ Fm_{sample} = \frac{R_{sample}\cdot Fm_{std}}{R_{std}}$$

If the expected fraction modern ($Fm_{std}$) of the standard is taken as a measured value with an error, rather than a constant or assigned ratio, normalization errors are propagated using this equation:

$$ \sigma_{Fm_{sample}} = Fm_{sample} \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2} +
\frac{\sigma_{Fm_{std}}^2}{Fm_{std}^2}} $$