### 2021-10-08 doc/AMS-15/AMS15_abstract.Rmd
Update abstract

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Josh Hlavenka
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### National Ocean Sciences AMS Facility, Woods Hole Oceanographic Institution, Woods Hole, MA, USA

We have developed a method for measurement of radiocarbon in carbonate minerals
as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
uses helium as a carrier gas to displace CO~2~ from sample vials to an open
split, where a glass capillary samples the mixture for delivery directly to the
HGIS. This method skips the gas transfer and quantification steps used in a
closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
mg carbonate can currently be measured. Results from measurements of consensus
standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
from a marine core (40-115 pMC) show that the method agrees well with
traditional AMS measurement of the same samples as graphite, and that the method
error is less than 1.5 pMC. We discuss advantages and disadvantages of
continuous flow sample introduction, and the effect of reduced precision on
calibrated age-depth models produced using gas-source data.