### 2021-10-08 doc/NAPB2021/NAPB2021.Rmd
Figures for NAPB2021

---
title: "HGIS Carbonate Results"
subtitle: "NAPB 2021"
author: 
  - "Brett Longworth"
date: '2021-10-08'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=6, fig.height=3, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

## Method

1. Acidify
  - ~30mg Carbonate 
  - produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials

![HGIS Figure](images/HGIS_open_split_drawing.svg)

## Data reduction

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

## Results: Agreement with Consensus

```{r}
# Read data. Don't use first carbonate test, smaller samples, or gas standards.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
mean_diff <- mean(cons_df$fm_diff)
sd_diff <- sd(cons_df$fm_diff)

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (cons - measured)")
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Pendleton Haiti Core Carbonates

### Read data, pull graphite results and sample depth, combine

```{r}
# Read raw and reduced data as produced by analyse_carbonates.R
data <- readRDS(here("data_analysed/carb_all.Rdata"))

# Read sample data
sample_data <- read_csv(here("data/carb_data.csv"))

# Select relevant reduced data, join with sample data
results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)

names <- results %>% 
  ungroup() %>% 
  filter(wheel == "USAMS062921") %>% 
  select(rec_num, sample_name)

graphite_results <- getRecOS(results$rec_num) %>% 
  filter(!is.na(reported)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
  mutate(method = "graphite") %>% 
  left_join(names, by = "rec_num")

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  ungroup()

depth <- read_csv(here("data/haiti_depth.csv"))

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

### Outlier

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr)
```

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite.

## Difference plot

```{r}
outlier_df <- combined_results %>% 
  filter(abs(fm_diff) > 0.02)

combined_results %>% 
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +#,
  #gghighlight::gghighlight(abs(fm_diff > 0.02)) +
  geom_pointrange(data = outlier_df,
                  aes(fm_corr, fm_diff, shape = method, color = method,
                      ymin = fm_diff - sig_fm_corr,
                      ymax = fm_diff + sig_fm_corr),
                  color = "grey",
                  size = 1) +#,
                  #position = position_dodge2(width = 0.02)) +
  #ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm") + 
  theme(legend.position = "none")
```

## Fm depth plot

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  labs(title = "Age vs depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)")
```

## BChron model

```{r}

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method)) %>% 
  select(id, depth, rc_age, sig_rc_age, method)

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gr <- cal_sub %>% 
  filter(method == "graphite") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gas <- cal_sub %>% 
  filter(method == "hgis") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

```

```{r}
plot(cal_dates)
```

```{r}
plot(cal_dates_gr)
```

```{r}
plot(cal_dates_gas)
```