### 2021-10-12 doc/AMS-15/AMS-15_poster.Rmd
Put together figures for AMS-15 poster

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Josh Hlavenka
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=6, fig.height=3, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)
library(gghighlight)
library(patchwork)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

Text and figures for AMS-15 poster.

# Abstract

We have developed a method for measurement of radiocarbon in carbonate minerals
as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
uses helium as a carrier gas to displace CO~2~ from sample vials to an open
split, where a glass capillary samples the mixture for delivery directly to the
HGIS. This method skips the gas transfer and quantification steps used in a
closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
mg carbonate can currently be measured. Results from measurements of consensus
standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
from a marine core (40-115 pMC) show that the method agrees well with
traditional AMS measurement of the same samples as graphite, and that the method
error is less than 1.5 pMC. We discuss advantages and disadvantages of
continuous flow sample introduction, and the effect of reduced precision on
calibrated age-depth models produced using gas-source data.

# Introduction

# Method

1. Acidify
  - ~30mg Carbonate 
  - produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials

![HGIS Figure](images/HGIS_open_split_drawing.svg)

## Data reduction

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

# Results

Data used for current, blanks, consensus

```{r}
# Read data. Don't use first carbonate test, smaller samples, or gas standards.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```
## Initial measurements

### Flow vs current

```{r cur-time-os, fig.height=7, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

# flow vs time subplot
flow_time <- data.frame(x = 0:125) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2"), size = 1.5) +
  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium"), size = 1.5) +
  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Gas flows to source",
       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
       y = "Gas flow (μL/min)") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.87, 0.65),
        legend.background = element_rect(fill = "white", color = "black"))

# current vs time subplot
cur_time <- ggplot(data, aes(time, he12C)) +
  geom_smooth(span = .4, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  xlim(0, 125) +
  labs(title = "Ion current",
       x = "Time (min)",
       y = "12C current (μA)") +
  theme_classic()

flow_time / cur_time
```

```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic()
```

compare to first tests

Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)

Also note current dependent ratio (blank) as a complicating factor

Lead into blank on next slide

### Blanks and blank correction

# Current independent blank

If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.

#### Linear fit used to correct:

$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$

#### Mean Fm of blanks: `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`

```{r blanks, fig.height=5, dev='svg'}
cons_df %>% 
  filter(fm_consensus < 0.1) %>% 
  ggplot(aes(he12C, norm_ratio)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
  geom_point(aes(color = wheel), size = 3) +
  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
  labs(title = "Blanks",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

## Results: Consensus standards

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
mean_diff <- mean(cons_df$fm_diff)
sd_diff <- sd(cons_df$fm_diff)

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (cons - measured)")
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Pendleton Haiti Core Carbonates

### Read data, pull graphite results and sample depth, combine

```{r}
# Read raw and reduced data as produced by analyse_carbonates.R
data <- readRDS(here("data_analysed/carb_all.Rdata"))

# Read sample data
sample_data <- read_csv(here("data/carb_data.csv"))

# Select relevant reduced data, join with sample data
results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)

names <- results %>% 
  ungroup() %>% 
  filter(wheel == "USAMS062921") %>% 
  select(rec_num, sample_name)

graphite_results <- getRecOS(results$rec_num) %>% 
  filter(!is.na(reported)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
  mutate(method = "graphite") %>% 
  left_join(names, by = "rec_num")

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  ungroup()

depth <- read_csv(here("data/haiti_depth.csv"))

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

### Outlier

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr)
```

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite.

## Difference plot

```{r}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

No outliers

```{r}
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

Highlight outliers

```{r}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  gghighlight(abs(fm_diff) > 0.02,
              #unhighlighted_params = list(colour = NULL),
              keep_scales = TRUE) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

All in one.

```{r}
outlier_df <- combined_results %>% 
  filter(abs(fm_diff) > 0.02)

ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1,
                  color = "grey") +#,
  gghighlight(abs(fm_diff) > 0.02,
              unhighlighted_params = list(colour = NULL),
              keep_scales = TRUE) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

## Fm depth plot

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  labs(title = "Age vs depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)")
```

## BChron model

```{r}

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method)) %>% 
  select(id, depth, rc_age, sig_rc_age, method)

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gr <- cal_sub %>% 
  filter(method == "graphite") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gas <- cal_sub %>% 
  filter(method == "hgis") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

```

```{r}
plot(cal_dates)
```

```{r}
plot(cal_dates_gr)
```

```{r}
plot(cal_dates_gas)
```