### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Add/edit text. Remove graphite value for hgis outlier.

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Josh Hlavenka
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=6, fig.height=3, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)
library(gghighlight)
library(patchwork)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

Text and figures for AMS-15 poster.

# Abstract

We have developed a method for measurement of radiocarbon in carbonate minerals
as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
uses helium as a carrier gas to displace CO~2~ from sample vials to an open
split, where a glass capillary samples the mixture for delivery directly to the
HGIS. This method skips the gas transfer and quantification steps used in a
closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
mg carbonate can currently be measured. Results from measurements of consensus
standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
from a marine core (40-115 pMC) show that the method agrees well with
traditional AMS measurement of the same samples as graphite, and that the method
error is less than 1.5 pMC. We discuss advantages and disadvantages of
continuous flow sample introduction, and the effect of reduced precision on
calibrated age-depth models produced using gas-source data.

# Introduction

# Method

![HGIS Photo](images/gas_control.jpg)

Our method uses an open split and glass capillary to control gas flow to the ion source. Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source. We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.

![HGIS Figure](images/HGIS_open_split_drawing.svg)

### Sample Preparation

1. 10-30mg Carbonate in septa-sealed 7mL vials
  - produces 2-6mL CO<sub>2</sub> @ ~1ATM
1. Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
1. Add helium to 1 ATM

![HGIS Photo](images/gilson.png)

### Measurement

1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Normalizing standards: live tank gas in vials
- Blanks: C-1 or TIRI-F in vials

## Data reduction

Normalization to standard, blank correction, and estimation of measurement error are done with the following steps:

1. Correct for fractionation using online AMS <sup>13</sup>C measurement.
1. Determine per-sample mean and error (repeatabilty and counting error).
1. Normalize to mean of Gas standard.
1. Blank correct: Blank subtraction using mean of carbonate blanks

Errors from normalization and blank correction are propagated at each step. The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.

Data reduction is done using an R package maintained on github (https://github.com/blongworth/hgis).

# Results

```{r}
# Read data. Don't use first carbonate test, smaller samples, or gas standards.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```

### Flow vs current

Introducing sample through an open split using helium displacement results in a decreasing flow of CO2 to the source over time. Measured ratio can depend on sample current, so CO2 flow rate must be kept in a range where ion currents are relatively constant with changing flow. We evaluated C- current over a range of CO2 flows and found that a range of flows produce a "flat top" in current. Less CO2 will starve the source of available material to dissociate and ionize, and too much CO2 also suppresses ionization.

```{r cur-time-os, fig.height=7, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

# flow vs time subplot
flow_time <- data.frame(x = 0:125) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2"), size = 1.5) +
  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium"), size = 1.5) +
  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Gas flows to source",
       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
       y = "Gas flow (μL/min)") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.87, 0.65),
        legend.background = element_rect(fill = "white", color = "black"))

# current vs time subplot
cur_time <- ggplot(data, aes(time, he12C)) +
  geom_smooth(span = .4, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  xlim(0, 125) +
  labs(title = "Ion current",
       x = "Time (min)",
       y = "12C current (μA)") +
  theme_classic()

flow_time / cur_time
```

```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
  geom_point(size = 3, color = "#0069b1") +
  labs(title = "Vial dilution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic()
```

### Blanks and blank correction

Measurements of live and dead gas over a range of C- currents show that there is a current-dependent component to the blank. We believe this is due to a combination of a constant amount of contaminant CO2 and a "target blank" current. This current dependent blank follows the constant contamination model and can be represented with a current balance between the sample and the constant blank.

While this current-dependent blank is important to consider for samples producing varying ion current, we simplified our assessment of blank by keeping sample currents as constant as possible for blanks, normalizing standards, and unknowns, which allows a simpler correction depending only on the Fm of the blank, standard, and unknown.

This leads to the following model for blank correction:

$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$

Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias. Our mean measured fraction modern of C-1 and TIRI-F carbonate blanks is `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`.

```{r blanks, fig.height=5, dev='svg'}
cons_df %>% 
  filter(fm_consensus < 0.1) %>% 
  ggplot(aes(he12C, norm_ratio)) +
  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
  geom_point(aes(color = wheel), size = 3) +
  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
  labs(title = "Blanks",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
mean_diff <- mean(cons_df$fm_diff)
sd_diff <- sd(cons_df$fm_diff)

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (cons - measured)")
```

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

## Marine Sediment Core Carbonates

To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method. We found the measurements to be in good agreement, with no significant bias and errors averaging XXX compared with XXX for graphite measurements.

```{r}
# Read raw and reduced data as produced by analyse_carbonates.R
data <- readRDS(here("data_analysed/carb_all.Rdata"))

# Read sample data
sample_data <- read_csv(here("data/carb_data.csv"))

# Select relevant reduced data, join with sample data
results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)

names <- results %>% 
  ungroup() %>% 
  filter(wheel == "USAMS062921") %>% 
  select(rec_num, sample_name)

graphite_results <- getRecOS(results$rec_num) %>% 
  filter(!is.na(reported)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
  mutate(method = "graphite") %>% 
  left_join(names, by = "rec_num")

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff_mean = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  left_join(select(graphite_results, rec_num, fm_corr_gr = fm_corr)) %>% 
  mutate(fm_diff = fm_corr - fm_corr_gr) %>% 
  ungroup()

depth <- read_csv(here("data/haiti_depth.csv"))

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

```{r, eval = FALSE}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

```{r}
# No outliers, vs depth
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

```{r, eval = FALSE}
# No outliers, vs mean_fm
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Mean Fm",
       y = "Fm - mean Fm")
```

```{r, eval=FALSE}
# Highlight outliers
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  gghighlight(abs(fm_diff) > 0.02,
              #unhighlighted_params = list(colour = NULL),
              keep_scales = TRUE) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

```{r eval=FALSE}
#All in one.

outlier_df <- combined_results %>% 
  filter(abs(fm_diff) > 0.02)

ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1,
                  color = "grey") +#,
  gghighlight(abs(fm_diff) > 0.02,
              unhighlighted_params = list(colour = NULL),
              keep_scales = TRUE) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

## Fm depth plot

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  labs(title = "Age vs depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)")
```

## BChron model

```{r}

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method),
         cal_curve = "intcal20") %>% 
  select(id, depth, rc_age, sig_rc_age, cal_curve, method) %>%
  add_row(id = "top-1", depth = 0, rc_age = 0, sig_rc_age = 1, cal_curve = "normal", method = "top")

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

cal_dates_gr <- cal_sub %>% 
  filter(method != "hgis") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

cal_dates_gas <- cal_sub %>% 
  filter(method != "graphite") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

```

```{r, eval=FALSE}
plot(cal_dates)
```

```{r}
plot(cal_dates_gr)
```

```{r}
plot(cal_dates_gas)
```

### Outlier

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite. This sample was dropped from the comparison. Not sure whether to mention this on poster.

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr, fm_diff)
```