### 2021-12-28 doc/AMS-15/AMS-15_paper.Rmd
add method paper outline

---
title: Analysis of Carbonates as CO2 using an open split interface to a hybrid sputter
  gas ion source
author: "Brett Longworth"
date: "7/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

NEC hybrid gas ion source

NEC inlets

MICADAS source and inlets

Comparison of open split to closed method

Pros/cons

# Methods

## Inlet description

## Sample preparation

## Data acquisition

## Data reduction

# Results

## System performance

### Currents

### Blanks

### Sample size

## Reference materials

## Haiti Core

# Discussion

Quantification of method

# Conclusion

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. Sample requirements, analysis time, precision, accuracy.