### 2022-03-25 R/HGIS_pendleton_all.Rmd
update library

---
title: "HGIS Pendleton Carbonates"
author: "Brett Longworth"
date: "2021-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(hgis)
library(rbacon)

options(digits = 4)
options(scipen = 3)
theme_set(theme_classic())
```

Data analysis for Haiti core carbonates from Simon Pendleton. Data reduction is handled by the `analyse_carbonates.R` script in the `HybridGIS_work` project using the method and functions from the `hgis` package.

- Data are normalized using a CO2 tank standard in sample vials.
- Method blank correction uses carbonate C-1 in vials and assumes that all samples have the same blank, including standards.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors

## Load results

```{r}
# Read raw and reduced data as produced by analyse_carbonates.R
data <- readRDS(here("data_analysed/carb_all.rds"))

# Read sample data
sample_data <- read_csv(here("data/carb_data.csv"))

# Select relevant raw data
raw <- map_dfr(data, 1) %>%  
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4)

# Select relevant reduced data, join with sample data
results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)

#write results to excel
#write.xlsx(results, here("data_analysed/haiti_carbonates.xlsx"))
```

### Currents and ratios vs time.

```{r, fig.height=10, fig.width = 10}
raw %>% 
  plot_hgis_time(norm_ratio, outlier = !.$ok_calc) +
  labs(title = "Sample ratio vs. time",
       y = "Fraction modern")
```

```{r, fig.height=10, fig.width = 10}
raw %>% 
  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
  labs(title = "Sample current vs time",
       x = "Time (s)",
       y = "Current (uA)")
```

## Results

```{r}
results %>% 
  mutate(he12C = he12C * 1E6) %>% 
  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, 
         `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
  arrange(Name) %>% 
  knitr::kable()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
results %>% 
  compare_replicates() %>% 
  select(Name, `Mean current (uA)` = he12C_mean, 
         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- results %>% 
  filter(!is.na(fm_consensus)) %>% 
  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amsdata::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(sample_name) %>% 
  select(Name = sample_name, 
         #`Mean current (uA)` = he12C_mean, 
         #`SD of current` = he12C_sd, 
         `Consensus Fm` = fm_consensus,
         Fm = fm_corr, 
         `Fm error` = sig_fm_corr,
         `Difference` = Fm_diff,
         Sigma = sigma)

knitr::kable(cons)
```

## Agreement with graphite

Agreement of Haiti core samples. Graphite samples were pretreated with HCl to leach 30% of the sample. HGIS samples not pretreated.

Load Simon's data and compare

```{r}
names <- results %>% 
  ungroup() %>% 
  filter(wheel == "USAMS062921") %>% 
  select(rec_num, sample_name)

graphite_results <- getRecOS(results$rec_num) %>% 
  filter(!is.na(reported)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
  mutate(method = "graphite") %>% 
  left_join(names, by = "rec_num")

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  ungroup()

combined_results %>% 
  select(sample_name, rec_num, method, fm_corr, sig_fm_corr) %>% 
  arrange(rec_num) %>% 
knitr::kable()
```

### Difference plot

```{r}
combined_results %>% 
  ggplot(aes(fm_corr, fm_diff, color = method, shape = method)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +#,
                  #position = position_dodge2(width = 0.02)) +
  #ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm") + theme_classic()
```

```{r}
combined_results %>% 
  ggplot(aes(rec_num, fm_corr, color = method, shape = method)) +
  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr),
                  size = .5,
                  position=position_dodge(width = .5)) +#,
                  #position = position_dodge2(width = 0.02)) +
  #ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       y = "Fraction modern") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  facet_wrap("Name", scales = "free")
```

### Outlier

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr)
```

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite.

## Age vs depth

Add depth data and radiocarbon age

```{r}
depth <- read_csv(here("data/haiti_depth.csv"))

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

Fm depth plot

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth() +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  ggtitle("Age vs depth for graphite and HGIS")
```

Mmmm.... Bacon

```{r}
# Bacon data format
# graphite
combined_results %>% 
  filter(method == "graphite",
         rc_age > 0) %>% 
  select(labID = Name, age = rc_age, error = sig_rc_age, depth) %>% 
  write_csv(here("Bacon_runs/haiti_graphite/haiti_graphite.csv"))

# hgis
combined_results %>% 
  filter(method == "hgis",
         sample_name != "6:7_39-40",
         !str_starts(sample_name, "HATC"),
         rc_age > 0) %>% 
  select(labID = Name, age = rc_age, error = sig_rc_age, depth) %>% 
  write_csv(here("Bacon_runs/haiti_hgis/haiti_hgis.csv"))
```

Trying Bchron here instead. Use is a bit friendlier in Rmarkdown.

```{r}
library(Bchron)

cal_sub <- combined_results %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method)) %>% 
  select(id, depth, rc_age, sig_rc_age)

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
plot(cal_dates)
```