### 2022-04-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Readthrough edits. Fix references.

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  pdf_document: default
  word_document: default
  html_document: default
bibliography: HGIS_paper.bib
csl: elsevier-without-titles.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message = FALSE}
library(tidyverse)
library(hgis)
library(here)
#library(errors)

#options(errors.notation = "plus-minus")

carb <- readRDS(here("data_analysed/carb_all.rds")) %>% 
  map_dfr(2)
```

# Abstract

We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1σ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.

# Introduction

Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials. A primary disadvantage of the method is its cost, which often limits the number of samples analyzed. This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination. One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite. Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly. The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a NEC 134-sample gas-capable MCSNICS source [@Xu2007] . We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.

Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018]. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary. Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split. This defines the delivery pressure as 1 ATM, and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement. The flow of CO~2~ to the source may be varied by controlling the ratio of CO~2~ to helium in the sample stream. Changing the ratio of CO~2~ to helium changes the partial pressure of CO~2~ and therefore the carbon flow rate to the source.

An open inlet has the advantage of simplicity; flow of gas to the source is preset by capillary parameters and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. Reduced system complexity also minimizes points of failure and eases automation and maintenance.

A disadvantage of the open split is that the flow of CO~2~ to the source varies with sample size (CO~2~:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO~2~ flow to a range that produces relatively constant C- ion current and stable ratios. In gas-source AMS, Roberts et al. @Roberts2013 used an open split to introduce sample gas to a ECR ion source and Salazar and Szidat @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.

The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007] works on the same principle as hybrid sources developed by ETH Zürich [@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004]. The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode. Sample gas is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species. Negative ions are extracted from the target, carbon species are selected, and interfering masses filtered out by the USAMS Tandetron beamline.

# Method

## Inlet description

The open split interface (Figure 1) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross fitting. Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in. PFTE tubing attached. The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split. The split has fairly low volume (< 1 mL). The system was operated manually during development, but may be mounted on the arm of a gasbench-style autosampler to enable automated operation.

```{r}
# Flow calc for CO2
co2flow <- flowcalc(100, 2.5E-5, 1.47E-5, 1.50)

# Flow calc for helium
heflow <- flowcalc(101, 2.5E-5, 1.96E-5, 1.50)
```

The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source was defined by 150 cm of 50 μm fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage. The section inside the cage was jacketed in 1/16 in. PFTE tubing to help protect against spark damage. The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.

## Sample preparation

Carbonate samples were weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W). Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added. The samples were allowed to react at 60 C for 4-6 h. Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum. Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].

Reference and process blank materials were not pretreated. Carbonate shell samples for the core comparison study were picked from a split core and cleaned by washing with distilled water and dried at 60 C. Shell samples were crushed and mixed, but not otherwise pretreated.

Standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system. Standards were adjusted to the same size as other samples by extracting an appropriate volume of CO~2~ from the full vial and replacing with helium using a syringe. Process blank samples were prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.

## Measurement

Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015]. The displacement helium was allowed to flow constantly during analysis, and the sample needle was inserted into a helium-filled vial between samples. Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA. The delivery needle was then transferred to the sample vial, which started sample delivery to the split and source. Currents were allowed to stabilize for 30 s , at which point the data acquisition was started. Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time. Normalizing standards, blanks, and secondary standards were measured identically to unknowns. Total time for a measurement was 20-25 min including target pre-cleaning and sample changes.

## Data reduction

After acquisition, data reduction, normalization and blank correction result in a final fraction modern with error. First, each 90 s block was corrected for fractionation using the AMS ^13/12^C measurement. The raw ^14/12^C ratio divided by the ^13/12^C ratio squared gives the corrected ^14/12^C ratio. Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were dropped at this point. Remaining runs of a sample were averaged and per-sample error was taken as the greater of counting error or the standard error of the measurements. The sample ratio was then normalized using the mean of gas standards measured during the run. The measurement error of the standards was taken as the greater of the standard error of all standards or the errors of the standards added in quadrature.

Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern [@donahue_isotope-ratio_1990; @roberts_14c_2019]. The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank. The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample. This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current. For all measurements used to assess the performance of the method and for the Haiti core comparison, sample, standard, and blank masses were kept constant. Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range. We discuss this constant contaminant blank in the results below. Data were reduced using a package written in R [@r_core_team_r_2022] which is published on github (<https://github.com/blongworth/hgis>).

# Results

## Source performance

```{r, message = FALSE}
carbcur <- carb %>%  
  compare_consensus() %>% 
  mutate(c12ua = he12C * 1E6)
samples <- read_csv(here("data/carb_data.csv")) %>% 
  select(wheel, pos, mass, he_added)
currents <- left_join(carbcur, samples, by = c("wheel", "pos")) %>% 
  filter(wheel != "USAMS020521")

curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("μA")
cursd <- sd(currents$c12ua)
```

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows. Best currents for the source was 10 μA ^12^C^-^, or roughly 1/10 of graphite target current. Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~. This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013]. Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean` ± `r cursd`. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].

Source currents are variable by day and by target under similar source, inlet, and sample conditions. Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium. Capillary clogging caused occasional flow issues, especially when sampling from vials containing acid.

A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium. We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2a) and comparing this to the composition of the gas delivered to the source. We estimated gas composition (Figure 2b) using a simple mixing model:

$$C(t) = e^{-\frac{r}{V}t}$$

Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial. Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate. The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement. Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et. al [@Xu2007] shows a somewhat broader peak with pure CO~2~. Broek et. al [@broek_optimization_nodate] found peak broadening with decreasing dilution ratio. While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.

Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 ml CO~2~ (equivalent to 0.8 mg C) remains in the sample vial. Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material. To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass. To ensure that samples started at or slightly above atmospheric pressure, we injected clean helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.

## Contaminant carbon

```{r}
# Fix to use current form of processed data
blanks <- carb %>% 
  filter(rec_num %in% c(2138, 83028),
         he12C > 2E-6,
         sample_type == "B")
blankmean <- mean(blanks$norm_ratio) %>% format(digits = 3)
blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 3)

# errors should produce an "all in one" for this, but it doesn't work with inline code
# meanblank <- as.errors(mean(blanks$norm_ratio), sd(blanks$norm_ratio))
nblank <- length(blanks$norm_ratio)
```

The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method. Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1σ uncertainty, N=`r nblank`). We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

HGIS cathodes produce measurable ion current, even with no gas flow. This current is initially \> 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes. For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline. This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020]. To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time. We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020]. The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method. Results from both tests are shown in Figure 3a. The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model. Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current. Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.

The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F^14^C = 0.82). A combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples may explain this difference[@sun_14_2020]. This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source. Given that this blank cathode current is a significant fraction of measured sample current, this current likely makes up a large fraction of the method blank.

While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size [@donahue_isotope-ratio_1990; @roberts_14c_2019]. If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current. Figure 3b shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current. Fraction modern of measured blanks does decrease with increasing current, but the decrease is small, less than F^14^C 0.01 over the range of currents measured. This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel. Constant current blank is likely the best model, but we simplify to a linear correction based on the measured ratio of process blanks and normalizing standards:

$$ R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}} $$

where $R_{samp}$ is the ratio of the unknown sample, $R_{meas}$ is the measured ratio, $R_{blank}$ is the ratio of the blank, and $R_{std}$ is the ratio of the normalizing standard. This mass-independent blank correction works well for our data since blanks vary by run and currents are relatively similar between blanks, unknowns and standards.

## Reference materials

```{r, message = FALSE}
#library(gt)
library(errors)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
options(errors.notation = "plus-minus")

# Format data
cons_df <- carb %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2"))

# Summary table
# cons_df %>% 
#   group_by(Name, fm_consensus) %>% 
#   summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
#                    list(mean = mean, sd = sd)),
#             N = n()) %>% 
#   ungroup() %>% 
#   arrange(fm_consensus) %>% 
#   mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
#           Name = case_when(Name == "TIRI-I" ~ "TIRI I",
#                            Name == "C-2" ~ "IAEA C2",
#                            Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
#   select(`Sample Name` = Name, 
#          `Reference F^14^C` = fm_consensus,
#          `HGIS F^14^C ± stdev` = fm_corr_err,
#          `Mean per-sample error` = sig_fm_corr_mean,
#          N) %>% 
#   knitr::kable(digits = 3,
#                caption = "Reference material precision and accuracy. Reference 
#                values are community consensus values for TIRI I and C2, and the 
#                NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
#                homogenized bivalve molusk reference. The mean of per-sample 
#                error estimates is included for comparison to the reproducibilty 
#                of the samples.")

  # gt() %>% 
  # tab_header(title = "Reference material precision and accuracy",
  #            subtitle = "Values in pMC, Errors are SD of measurements") %>% 
  # fmt_number(4, 
  #            decimals = 3) %>% 
  # cols_align("right", 3)

# Consensus summary for inline data
cons_sum <- cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff),
                   list(mean = mean, sd = sd)),
            N = n()) 
```

Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements. We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve molusk REF. Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information. Measurement, data reduction, and blank correction was performed as described above.

Reference materials run with the hybrid gas method were in good agreement with consensus values. Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate. The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1σ precision of the method. The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above. The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.

The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4). Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1σ uncertainties than would be expected statistically.

## Carbonate macrofossils for age-depth model

```{r core data, message = FALSE}
core <- read_csv(here("data_analysed/haiti_combined.csv"))

# Remove outlier: HGIS sample leaked
# Remove HATC duplicates
cr_no <- core %>% 
  filter(rec_num != 171996,
         !str_starts(sample_name, "HATC"))

# Data for duplicate discussion
dups <- core %>% 
  filter(method == "hgis") %>% 
  group_by(rec_num) %>% 
         filter(n() > 1) %>% 
  ungroup()

# pick single core run for fairness?
core_summary <- cr_no %>% 
  group_by(method) %>% 
  #filter(method == "hgis") %>% 
  summarize(across(c(sig_fm_corr, fm_diff),
                   list(mean = mean, sd = sd)),
            N = n())
```

A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser_multi-proxy_2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application. Sediment cores were collected from Baie des Baradères, Haiti during a scientific cruise in May 2016 lead by Dr. Jeffrey Donnelly (WHOI). Coring was performed with a Rossfelder vibracorer from a 20-foot pontoon boat. The cores were collected from a coastal karst basin (CKB) at vertical depth ~11.7 m and coordinates 18.514° N, 73.629° W. Core HATC1 D1 is 145 cm long and was collected in a 3-inch poly tube as a control directly beside core HATC1 D2, which is a continuous 896 cm core collected in a 3-inch aluminum tube.

Carbonate materials such as corals and bivalve shells were selected throughout the core for the age-depth model, processed and measured as described above. Parallel graphite measurements were made following NOSAMS methods for acidification, graphitization, and measurement. One HGIS sample vial was noted to be at atmospheric pressure when the 16 mg sample should have produced only 0.5 atmosphere of CO~2~. The HGIS measurement for this sample was 0.053 F^14^C higher than the measurement as graphite. We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.

We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 5a). Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C. Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite. Three samples were analysed by HGIS in duplicate. Two of the three pairs agreed within 1σ and all three agreed within 2σ. There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value. The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.

```{r}
# Agreement within uncertainty of paired ages?
# Mean and sd of difference of calibrated ages?
```

Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates. We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models. We applied a correction for the regional marine reservoir effect with a delta R of -216 and uncertainty 92 based on the oxcal model with 10 reservoir points. Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020. The NH zone 2 curve from Hua et. al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction. The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.

# Discussion

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. This system produces results that agree with graphite measurements within the approximately 1% method error from 8-20 mg of carbonate. Sample preparation takes about 10 min. of person time and 6 h of automated processing and reaction time, and 20 min. of measurement time per sample. The simplicity of of this approach reduces costs, improves reliability, and reduces system maintenance.

With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, i.e. triplicate measurement or more core depths. This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.

The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018]. The lower time and cost of the hybrid method allows analysis of more samples. This may produce more accurate age models, despite reduced precision of individual measurements.

# Acknowledgements

This work was supported by the NOSAMS NSF cooperative agreement 1755125. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, and T. Broek for discussions on gas handling and gas ion sourcery; and to J. Gunnard for improvements to the text.

# Tables

```{r message = FALSE}
cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup() %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± stdev` = fm_corr_err,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
  knitr::kable(digits = 3,
               caption = "Table 1. Reference material precision and accuracy. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
               homogenized bivalve molusk reference. The mean of per-sample 
               error estimates is included for comparison to the reproducibilty 
               of the samples.")

```

# Figures

![Figure 1](images/HGIS_open_split_drawing.svg)

![Figure 2](figures/fig2_current_flow.svg)

![Figure 3](figures/fig3_blanks.svg)

![Figure 4](figures/fig4_consensus_samples.svg)

![Figure 5](figures/fig5_core_carbonates.svg)

# Figure captions

Figure 1. Schematic of the open split inlet. Two valco valves provide the ability to admit helium and live or dead tank standards. The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.

Figure 2. Performance of the NOSAMS MCSNICS hybrid gas ion source. (a) Measured ^12^C^-^ ion current during displacement of a pure CO~2~ vial sample with helium. (b) Modeled flows of helium and CO~2~ to the ion source during the same dilution experiment. (c) Measured ion current for modeled CO~2~ flow. The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.

Figure 3. Assessment of blanks. (a) Current dependent blank model and data. F^14^C and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the F^14^C and current of the constant current blank. Direct determination of F^14^C and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.

Figure 4. Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.

Figure 5. Comparison of Haiti core samples measured as graphite and via HGIS. (a) Difference between HGIS and graphite measurements of the same materials. Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements. Samples are offset on the horizontal axis for clarity. Error bars are measurement error for each sample. Graphite measurements are included to allow comparison of error estimates. The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).

# References