### 2022-04-14 doc/AMS-15/AMS-15_paper/tables/table_s2_haiti_core.Rmd

```{r}
library(tidyverse)
library(here)
library(hgis)
library(errors)

options(errors.notation = "plus-minus")
options(scipen = 999)

# Read data. Don't use first carbonate test, smaller samples, or gas standards.
core <- read_csv(here("data_analysed/haiti_combined.csv")) %>% 
  select(sample_name, depth, method, fm_corr, sig_fm_corr)

dups <- core %>% 
  filter(str_starts(sample_name, "HATC")) %>% 
  select(sample_name, depth, fm_corr_hgis = fm_corr, sig_fm_corr_hgis = sig_fm_corr) %>% 
  mutate(fm_corr_graphite = NA, sig_fm_corr_graphite = NA)

core_long <- core %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  pivot_wider(names_from = method, values_from = c(fm_corr, sig_fm_corr)) %>% 
  bind_rows(dups) %>% 
  arrange(depth)

core_long %>% 
  write_csv(here("data_analysed/haiti_summary.csv"))

core_long %>% 
  mutate (hgis = set_errors(fm_corr_hgis, sig_fm_corr_hgis),
          graphite = set_errors(fm_corr_graphite, sig_fm_corr_graphite)) %>% 
  select(`Sample Name` = sample_name, 
         `Core Depth` = depth,
         `HGIS F^14^C ± 1$\\sigma$` = hgis,
         `graphite F^14^C ± 1$\\sigma$` = graphite) %>% 
  knitr::kable(digits = c(3, 4, 4, 4),
               caption = "\\label{tabS2}HGIS and graphite measurements of Haiti core carbonates.
                          HGIS replicates are listed without a corresponding graphite value.")
```